---
title: "Diversity and Plant response to microbes (PRM)"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
***

To make figures and statistics for cross examining PRM and diversity we obtained data generated from our "Diversity Analysis.Rmd" script (Providing us diverersity values for each sample)  and "res_tuber_analysis2.Rmd" (Providing us PRM values)

#Step 1: import relevant datasets:

```{r}
library(tidyverse)
res_data = read.csv("residual_tubers.csv", header = T)
div_data = read.csv("metadata.csv", header = T)

#Now we need to make sure header labels match for each dataset:
labels(res_data)
labels(div_data)

#need to rename "Variety2" to "Variety" in res_data dataframe:
res_data = res_data %>%
  rename(Variety = variety2) %>%
  rename(Soil_type = soil)
labels(res_data)

```


##Timepoint 2:
```{r}
#metadata2 <- read.csv(file = "metadata2.csv", header = T)

ghsoilspecies_16S_div = div_data %>%
  filter(timepoint != "16S") %>%
  filter(Soil_type != "Sterile") %>%
  group_by(Nutrient,Soil_type,Variety,Species2) %>%
  summarize(shannon = mean(diversity))

labels(ghsoilspecies_16S_div)
labels(res_data)

ghsoilspecies_16S_div2 <- left_join(res_tubers2,ghsoilspecies_16S_div, by = c("Nutrient", "Soil_type","Variety"))

ghsoilspecies_16S_div2 <- subset(ghsoilspecies_16S_div2, shannon != "NA")

```

**When calculating shannon index make sure to use 

Testing plant microbial response determined by shannon diveristy index and Plant Clade @ low nutrient condition:

##Timepoint 2

@Low nutrient conditions at timepoint2: *Use for Table 4*
```{r}
ghsoilspecies_16S_div2$agerankcat <- as.factor(ghsoilspecies_16S_div2$agerankcat)

DivvsRes <- lmer(res ~ shannon * agerankcat + (1|Variety), data = subset(ghsoilspecies_16S_div2, Nutrient == "Low") )
summary(DivvsRes)
anova(DivvsRes)

```


Testing plant microbial response determined by shannon diveristy index and Plant Clade @ high nutrient condition:
```{r}

DivvsRes <- lmer(res ~ shannon * agerankcat + (1|Variety), data = subset(ghsoilspecies_16S_div2, Nutrient == "High") )
summary(DivvsRes)
anova(DivvsRes)

```


```{r}
#Low and timepoint 2 Nutrient Conditions:

div_resL <- subset(ghsoilspecies_16S_div2, Nutrient == "Low")
div_resL <- rename(div_resL, Clade = Species2) 


ggplot(div_resL)+
  aes(x = shannon, y = res, color = Clade)+
  geom_point(aes(shape = Clade), size = 3)+
  scale_color_manual(values = c("Black","#117733","Orange", "#0072B2"))+
  labs(x = "Bacterial Diversity", y = "Plant Response to Microbes")+
  theme_classic()+
  theme(axis.title.x = element_text(face="bold", size=15))+
  theme(axis.title.y = element_text(face="bold", size=15))


#High Nutrient and timepoint 2 Conditions
div_resH <- subset(ghsoilspecies_16S_div2, Nutrient == "High")
div_resH <- rename(div_resH, Clade = Species2) 


ggplot(div_resH)+
  aes(x = shannon, y = res, color = Clade)+
  geom_point(aes(shape = Clade), size = 3)+
  scale_color_manual(values = c("Black","#117733","Orange", "#0072B2"))+
  labs(x = "Bacterial Diversity", y = "Plant Response to Microbes")+
  theme_classic()+
  theme(axis.title.x = element_text(face="bold", size=15))+
  theme(axis.title.y = element_text(face="bold", size=15))

```



##Timepoint 1:

We want to now look at timepoint 1 (Week 4) and how it may explain tuber yield
```{r}
library(tidyverse)
ghsoilspecies_16S_div = metadata %>%
  filter(timepoint != "16S2") %>%
  filter(Soil_type != "Sterile") %>%
  group_by(Nutrient,Soil_type,Variety,Species2) %>%
  summarize(shannon = mean(diversity))

ghsoilspecies_16S_div3 <- left_join(ghsoilspecies_16S_div,res_tubers2, by = c("Nutrient", "Soil_type","Variety"))

ghsoilspecies_16S_div3 <- subset(ghsoilspecies_16S_div3, shannon != "NA")
ghsoilspecies_16S_div3$agerankcat <- as.factor(ghsoilspecies_16S_div3$agerankcat)
```


```{r}

DivvsResL <- lmer(res ~ shannon * agerankcat + (1|Variety), data = subset(ghsoilspecies_16S_div3, Nutrient == "Low") )
summary(DivvsResL)
anova(DivvsResL)


DivvsResH <- lmer(res ~ shannon * agerankcat + (1|Variety), data = subset(ghsoilspecies_16S_div3, Nutrient == "High") )
summary(DivvsResH)
anova(DivvsResH)

```

*Plot Plant microbial response vs Shannon Diveristy (Figure 5A-B)*

```{r}
#Low and timepoint 2 Nutrient Conditions:

div_resL <- subset(ghsoilspecies_16S_div3, Nutrient == "Low")
div_resL <- rename(div_resL, Clade = Species2) 

ggplot(div_resL)+
  aes(x = shannon, y = res, color = Clade)+
  geom_point(aes(shape = Clade), size = 3)+
  scale_color_manual(values = c("Black","#117733","Orange", "#0072B2"))+
  theme_classic()+
  labs(x = "Bacterial Diversity", y = "Plant Response to Microbes")+
  theme(axis.title.x = element_text(face="bold", size=15))+
  theme(axis.title.y = element_text(face="bold", size=15))


#High Nutrient and timepoint 2 Conditions

div_resH <- subset(ghsoilspecies_16S_div3, Nutrient == "High")
div_resH <- rename(div_resH, Clade = Species2) 

ggplot(div_resH)+
  aes(x = shannon, y = res, color = Clade)+
  geom_point(aes(shape = Clade), size = 3)+
  scale_color_manual(values = c("Black","#117733","Orange", "#0072B2"))+
  theme_classic()+
  labs(x = "Bacterial Diversity", y = "Plant Response to Microbes")+
  theme(axis.title.x = element_text(face="bold", size=15))+
  theme(axis.title.y = element_text(face="bold", size=15))


```





