---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---



```{r, echo=FALSE}
library(tidyverse)
library(vegan)
library(ggplot2)

gh_soil_16S_rdp = read.csv("GHexp17max-Rick1.csv",header = T)
labels(gh_soil_16S_rdp)
str(gh_soil_16S_rdp)
rarecurve(gh_soil_16S_rdp[28:1505], col = "blue")

table2 <- gh_soil_16S_rdp[29:1505]

```


We see that alot of our samples start to plataeu around ~300 reads. And so we will now proceed to do some our quality control step. 

Note:In excel we added the row "Total" which counts the number of ASVs per sample. Doing so we can do the following quality control step: 


```{r}
ghsoilspecies_16S_edited <- subset(gh_soil_16S_rdp,total > 300 & Soil_type != "Sterile")
ghsoilspecies_16S_edited <- subset(ghsoilspecies_16S_edited, Soil_type != "Sterile")
write.csv(ghsoilspecies_16S_edited, file = "ghsoil16S.csv")

```



Now we need to normalize our counts:

```{r, echo=FALSE}

#Normalize asv data with decostand function:
library(vegan)
ghsoilspecies_16S_std = decostand(ghsoilspecies_16S_edited[,28:1505], method = "total")
ghsoil16S = cbind(ghsoilspecies_16S_edited[,1:26],ghsoilspecies_16S_std)
write.csv(ghsoil16S, file = "ghsoil16S_2.csv")

#Now we will ordinate using NMDS method: 
set.seed(1)
ghsoilmds <- metaMDS(ghsoil16S[27:1504], distance = "bray", k = 2)
save(ghsoilmds, file = "ghsoilmds.rda")

plot(ghsoilmds, display = "species", type= "n")
points(ghsoilmds,display = "sites", col = as.numeric(ghsoil16S$Species2), bg = ghsoil16S$Species2, pch = as.numeric(ghsoil16S$timepoint)+ 14)
ordiellipse(ghsoilmds, ghsoil16S$Species2, label = F, draw = "polygon", col = c("Black","Blue","Red", "Purple", alpha = 70))
legend("topright", legend = levels(ghsoilspecies_low$Species2), bty = "n", col = c("Black","Blue","Red", "Purple"), pch = as.numeric(ghsoil16S$Species_timepoint)+ 14)
#ghsoil16S <- read.csv("ghsoil16S_1.csv",header = T)
```
**no convergence** why is this happening all of the sudden....


Stress value is 0.23 which is okay. Its not ideal. 
Now we proceed to plot our NMDS plot: 

```{r, echo=FALSE}
summary(ghsoilmds)
plot(ghsoilmds, display = "species", type= "n")
points(ghsoilmds,display = "sites", col = as.numeric(ghsoil16S$Species2))
ordiellipse(ghsoilmds, ghsoil16S$Species, label = TRUE)
plot(ghsoilmds, display = "species", type= "n")
points(ghsoilmds,display = "sites", col = as.numeric(ghsoil16S$Nutrient))
ordiellipse(ghsoilmds, ghsoil16S$Nutrient, label = TRUE)
plot(ghsoilmds, display = "species", type= "n")
points(ghsoilmds,display = "sites", col = as.numeric(ghsoil16S$Soil_type))
ordiellipse(ghsoilmds, ghsoil16S$Soil_type, label = TRUE)
plot(ghsoilmds, display = "species", type= "n")
points(ghsoilmds,display = "sites", col = as.numeric(ghsoil16S$timepoint))
ordiellipse(ghsoilmds, ghsoil16S$timepoint, label = TRUE)
```
This plot is of everything! Made a new data set! called meta_coords which contains both NMDS coordinates and the associated metadata. 

#Community Structure Using NMDS:


High vs Low Nutrient Conditions:

*Figure 2*
```{r}
ghsoilspecies_16S_High <- subset(ghsoil16S, Nutrient == "High")
ghsoilmds = metaMDS(ghsoilspecies_16S_High[,27:1504])
plot(ghsoilmds, display = "sites", type= "n")
  points(ghsoilmds,display = "sites", col = as.numeric(ghsoilspecies_16S_High$Species2), bg = ghsoilspecies_16S_High$Species2, pch = as.numeric(ghsoilspecies_16S_High$timepoint)+ 14)
  ordiellipse(ghsoilmds, ghsoilspecies_16S_High$Species2, label = T, draw = "polygon", col = c("Black","Blue","Red", "Purple", alpha = 70))
  legend("topright", legend = levels(ghsoilspecies_16S_High$Species2), bty = "n", col = c("Black","Blue","Red", "Purple"), pch = as.numeric(ghsoilspecies_16S_High$Species2)+ 14)
```

-------------------------------------------------------------------------------

*Figure 2A and 2B*

```{r}
ghsoilspecies_16S_Low <- subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S2")

#labels(ghsoilspecies_16S_Low)
#will tell us what columns to run metaMDS function with

ghsoilmds = metaMDS(ghsoilspecies_16S_Low[27:1504])
plot(ghsoilmds, display = "sites", type= "n")
  points(ghsoilmds,display = "sites", col = as.numeric(ghsoilspecies_16S_Low$Species2), pch = as.numeric(ghsoilspecies_16S_Low$timepoint)+14)
  ordiellipse(ghsoilmds, ghsoilspecies_16S_Low$Species2, col = 1:4, border = 1:4, alpha = 63, label = F)
  legend(x = "bottomleft", legend = c(levels(ghsoilspecies_16S_Low$Species)), col = 1:4, pch=20)

```

Species effect as changed, Nutient effect kicks in! You see that everything else is statistically significantly indifferent than Modern in timepoint 2 in low nutrient conditions. We see a slight species effect at 4 weeks in Hancock soil ($p = 0.050$) but not so much Virgin Soil ($p = 0.79$).


```{r}
ghsoilspecies_16S_Low <- subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S")

#labels(ghsoilspecies_16S_Low)
#will tell us what columns to run metaMDS function with

ghsoilmds = metaMDS(ghsoilspecies_16S_Low[27:1504])
plot(ghsoilmds, display = "sites", type= "n")
  points(ghsoilmds,display = "sites", col = as.numeric(ghsoilspecies_16S_Low$Species2), pch = as.numeric(ghsoilspecies_16S_Low$timepoint)+14)
  ordiellipse(ghsoilmds, ghsoilspecies_16S_Low$Species2, col = 1:4, border = 1:4, alpha = 63, label = F)
  legend(x = "bottomleft", legend = c(levels(ghsoilspecies_16S_Low$Species)), col = 1:4, pch=20)

```


##Want to see nutrient at low levels at both timepoints

*Figure 2A and Figure 2B for paper*
```{r}
ghsoilspecies_16S <- subset(ghsoil16S, Nutrient == "Low")
ghsoilmds = metaMDS(ghsoilspecies_16S[,27:1504])

colors = c("Black","#117733","Orange", "#0072B2")
colors1 = as.numeric(ghsoilspecies_16S$Species2)
colors1 = replace(colors1, colors1 == 1, "Black")
colors1 = replace(colors1, colors1 == 2,"#117733")
colors1 = replace(colors1,colors1 == 3,"Orange")
colors1 = replace(colors1,colors1 == 4, "#0072B2")


plot(ghsoilmds, display = "sites", type= "n")
points(ghsoilmds,display = "sites", col = colors1, pch = as.numeric(ghsoilspecies_16S$timepoint)+ 14)
ordiellipse(ghsoilmds, ghsoilspecies_16S$Species_timepoint, label = F, draw = "polygon", col = colors, alpha = 70, kind = "se", conf = 0.95)
legend("topright", legend = c("Andigenum", "Chiletanum", "Modern", "S. berthaulti"), bty = "n", col = colors, pch = 20)
legend(x = "bottomright", legend = c("Week 4","Week 12"), pch=c(15,16),bty = "n")

#Things done: 1) Fixed results so that they match the statistic tables! 2) Fixed legends to tell reader what is what. 3) Fixed the legends from Max Speak to normal 
#Things to be done: 1) Need to fix the color scheme, so that it matches other figures. 2) Need to fix maxspeak on the tables

```

*Figure 2B for paper*
```{r}

colors = c("Black","#117733","Orange", "#0072B2")
colors1 = as.numeric(gh_soil_16S_low$Species2)
colors1 = replace(colors1, colors1 == 1, "Black")
colors1 = replace(colors1, colors1 == 2,"#117733")
colors1 = replace(colors1,colors1 == 3,"Orange")
colors1 = replace(colors1,colors1 == 4, "#0072B2")


ghsoilspecies_16S2 <- subset(ghsoil16S, Nutrient == "High")
ghsoilmds = metaMDS(ghsoilspecies_16S2[,27:1504])
plot(ghsoilmds, display = "sites", type= "n")
points(ghsoilmds,display = "sites", col = colors, bg = colors, pch = as.numeric(ghsoilspecies_16S2$timepoint)+ 14)
ordiellipse(ghsoilmds, ghsoilspecies_16S2$Species_timepoint, label = F, draw = "polygon", col = colors, alpha = 70)
legend("topright", legend = c("Andigenum", "Chiletanum", "Modern", "S. berthaulti"), bty = "n", col = colors , pch = 20)
legend(x = "bottomright", legend = c("Week 4","Week 12"), bty = "n", pch=c(15,16))
```



##Full Model of Community structure:

```{r}
library(vegan)
fit_gh16S <- adonis(ghsoil16S[27:1504] ~ timepoint * Nutrient * Soil_type * Species2, data = ghsoil16S, permutations = 999)
fit_gh16S$aov.tab
```

We see that Species has a not a strong effect $p=0.121$
We also see that timepoint, Nutrient, and soil type has a very strong effect $p < 0.05$


##At timepoint and low nutrient conditions:

low nutrient conditions:
*Table 2A (Used in paper)*
```{r}
ghsoilspecies_16S2 = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S2")
fit_gh16S <- adonis(ghsoilspecies_16S2[27:1504] ~ Soil_type + Species2, data = ghsoilspecies_16S2, permutations = 999)
fit_gh16S

ghsoilspecies_16S2 = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S")
fit_gh16S <- adonis(ghsoilspecies_16S2[27:1504] ~ Soil_type + Species2, data = ghsoilspecies_16S2, permutations = 999)
fit_gh16S
```
Species effect ($p = 0.110$) in low nutrient conditions may exist. But we see that soil inocula effect is present at timepoint 1 but no longer at timepoint 2. 

*Table 2B (Used in paper)*
```{r}
ghsoilspecies_16S2 = subset(ghsoil16S, Nutrient == "High" & timepoint == "16S2")
fit_gh16S <- adonis(ghsoilspecies_16S2[27:1504] ~ Soil_type + Species2, data = ghsoilspecies_16S2, permutations = 999)
fit_gh16S

ghsoilspecies_16S = subset(ghsoil16S, Nutrient == "High" & timepoint == "16S")
fit_gh16S <- adonis(ghsoilspecies_16S[27:1504] ~ Soil_type + Species2, data = ghsoilspecies_16S, permutations = 999)
fit_gh16S

```
Species effect is ($p = 0.552$) and doesn't exist at timepoint 1. However, it does exist at timepoint 2 tho which is interesting ($p = 0.038$). 


##Comparison within timepoints: 

```{r}
ghsoilspecies_16S1 = subset(ghsoil16S, timepoint == "16S")
fit_gh16S <- adonis(ghsoilspecies_16S1[27:1504] ~ (Nutrient + Soil_type) * Species2, data = ghsoilspecies_16S1, permutations = 999)
fit_gh16S
```
See that in timepoint 1 (Week4) one we see soil_type and species affecting community composition. 


```{r}
ghsoilspecies_16S1 = subset(ghsoil16S, timepoint == "16S2")
fit_gh16S <- adonis(ghsoilspecies_16S1[27:1504] ~ (Nutrient + Soil_type) * Species2, data = ghsoilspecies_16S1, permutations = 999)
fit_gh16S
```
See no differences after timepoint 2 in the experimental variables that are leading to changes in community structure. Initially start out different and then start to converge. 


##Pairwise statistics for microbial communities (Table 3) 


```{r}
#Global pairwise analysis of 16S data:

#ghsoilspecies_16S_edited is not standardized! 

ghsoilspecies_16S <- subset(ghsoil16S, Nutrient == "Low")


library(pairwiseAdonis)
gh_soil_16S = subset(ghsoil16S, timepoint == "16S")
distance_matrix = vegdist(gh_soil_16S[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_low = merged_col <- c(gh_soil_16S[,0:23],distance)
bray_low <- as.data.frame(bray_low)
low = pairwise.adonis(bray_low[,24:88],bray_low$Species2)
low

gh_soil_16S = subset(ghsoil16S, timepoint == "16S2")
distance_matrix = vegdist(gh_soil_16S[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_low = merged_col <- c(gh_soil_16S[,0:23],distance)
bray_low <- as.data.frame(bray_low)
low = pairwise.adonis(bray_low[,24:86],bray_low$Species2)
low
```



```{r}
#pairwise statistic for timepoint 2 and low nutrient:

library(pairwiseAdonis)
gh_soil_16S_low = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S2")
distance_matrix = vegdist(gh_soil_16S_low[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_low = merged_col <- c(gh_soil_16S_low[,0:23],distance)
bray_low <- as.data.frame(bray_low)
low = pairwise.adonis(bray_low[,24:58],bray_low$Species2)
low
```
Only see differences in microbial communities between E and SB after 12 weeks...


```{r}
gh_soil_16S_low = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S")
distance_matrix = vegdist(gh_soil_16S_low[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_low = merged_col <- c(gh_soil_16S_low[,0:23],distance)
bray_low <- as.data.frame(bray_low)
low = pairwise.adonis(bray_low[,24:57],bray_low$Species2)
low
```
We do see differences in microbial communities in timepoint 1


```{r}
#pairwise statistic for timepoint 2 and high nutrient

library(pairwiseAdonis)
gh_soil_16S_high = subset(ghsoil16S, Nutrient == "High" & timepoint == "16S2")
distance_matrix = vegdist(gh_soil_16S_high[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_high = merged_col <- c(gh_soil_16S_high[,0:23],distance)
bray_high <- as.data.frame(bray_high)
high = pairwise.adonis(bray_high[,24:51],bray_high$Species2)
high
```



```{r}
gh_soil_16S_high = subset(ghsoil16S, Nutrient == "High" & timepoint == "16S")
distance_matrix = vegdist(gh_soil_16S_high[27:1504], method = "bray")
distance_matrix = as.matrix(distance_matrix)
distance = as.data.frame(distance_matrix)
bray_high = merged_col <- c(gh_soil_16S_high[,0:23],distance)
bray_high <- as.data.frame(bray_high)
high = pairwise.adonis(bray_high[,24:54],bray_high$Species2)
high
```
There's no difference in microbial communities found in High nutrient conditions


***

*GRAPHING Figure 2A,2B,2C,2D*

```{r}

gh_soil_16S_low = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S2")
gh_soil_16S_low_std = decostand(gh_soil_16S_low[27:1504], method = "total")
pairwise_mds <- metaMDS(gh_soil_16S_low_std)


colors = c("Black","#117733","Orange", "#0072B2")
colors1 = as.numeric(gh_soil_16S_low$Species2)
colors1 = replace(colors1, colors1 == 1, "Black")
colors1 = replace(colors1, colors1 == 2,"#117733")
colors1 = replace(colors1,colors1 == 3,"Orange")
colors1 = replace(colors1,colors1 == 4, "#0072B2")



plot(pairwise_mds, display = "sites", type = "n")
points(pairwise_mds,display = "sites", col = colors1, pch = 16)
ordiellipse(pairwise_mds, gh_soil_16S_low$Species2, draw = "polygon", col = colors, border = colors, alpha = 70, label = F, kind = "se", conf = 0.95)
legend("bottomleft", legend = c("Andigenum", "Chiletanum", "Modern", "S. berthaulti"), bty = "n", col = colors , pch = 20)

```



```{r}

gh_soil_16S_low = subset(ghsoil16S, Nutrient == "Low" & timepoint == "16S")
pairwise_mds <- metaMDS(gh_soil_16S_low[27:1504])

plot(pairwise_mds, display = "sites", type = "n")
points(pairwise_mds,display = "sites", col = colors1, pch = 16)
ordiellipse(pairwise_mds, gh_soil_16S_low$Species2, draw = "polygon", col = colors, border = colors, alpha = 70, label = T, kind = "se", conf = 0.95)
legend("topright", legend = levels(ghsoilspecies_low$Species2), bty = "n", col = colors)
```

```{r}
gh_soil_16S_high = subset(ghsoil16S, Nutrient == "High" & timepoint == "16S")
pairwise_mds <- metaMDS(gh_soil_16S_high[27:1504])

plot(pairwise_mds, display = "sites", type = "n")
points(pairwise_mds,display = "sites", col = colors1, pch = as.numeric(gh_soil_16S_high$timepoint)+14)
ordiellipse(pairwise_mds, gh_soil_16S_high$Species2, col = colors, border = colors, draw = "polygon", alpha = 70, label = F)
```


```{r}
gh_soil_16S_high = subset(ghsoilspecies_16S_edited, Nutrient == "High" & timepoint == "16S2")
pairwise_mds <- metaMDS(gh_soil_16S_high[27:1504])
plot(pairwise_mds, display = "sites", type = "n")
points(pairwise_mds,display = "sites", col = as.numeric(gh_soil_16S_high$Species2), pch = as.numeric(gh_soil_16S_high$timepoint)+14)
ordiellipse(pairwise_mds, gh_soil_16S_high$Species2, draw = "polygon" , col = colors, border = 1:4, alpha = 70, label = F)
```


***


DATA analysis and generation for Table 4

***


#Let's try to correlate diversity and tuber yield**

*Use for Table 4*
```{r}
library(lme4)
library(lmerTest)

div_test = lmer(Tubermass2 ~ (Soil_type+diversity)*Species.cat + (1|Variety), data = subset(metadata, timepoint == "16S2" & Nutrient == "Low"))
summary(div_test)
anova(div_test)


```



*Use for Paper Table 4*
```{r}

metadata$timepoint = as.factor(metadata$timepoint)

div_test <- lmer(Tubermass2 ~ (Soil_type+diversity)*Species.cat + (1|Variety), data = subset(metadata, timepoint == "16S2" & Nutrient == "Low"))

summary(div_test)
anova(div_test)

```
At both timepoints we see a effect of diversity on tuber yield variation ($0.0515562$). In terms of the mean tubermass produced. In Low Nutrient conditions diveristy does play some role in determining tuber yield as well. 



Pull S.E from Lme object!(Use emmeans)
*https://cran.r-project.org/web/packages/emmeans/vignettes/comparisons.html



```{r}
metadata$timepoint = as.factor(metadata$timepoint)

div_testA <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "A"))

summary(div_testA)
anova(div_testA)
```
Low Nutrient conditions has a effect on the average tubermass for Andigenum. But diveristy of microbes still has not much of a effect. 


```{r}
metadata$timepoint = as.factor(metadata$timepoint)

div_testC <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity*timepoint + (1|Variety), data = subset(metadata, Species == "C"))

summary(div_testC)
anova(div_testC)
```
For Chiletnaum we see that diveristy affects tuber yield production ($p=0.0148$, t-test). We see that the with increasing diversity we see a decrease in tuber yield. In addition, we see variation in tubermass is partially explained by a Nutrient:Diversity interaction effect (ANOVA table, $p=0.059412$). This could be obseved in that diversity has a positive effect on tubermass production in low nutrient conditions (Summary table). 


```{r}
metadata$timepoint = as.factor(metadata$timepoint)

div_testSB <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "SB"))

summary(div_testSB)
anova(div_testSB)
```
Weird result. Everything effects S. berthauti. 

```{r}
metadata$timepoint = as.factor(metadata$timepoint)

div_testE1 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "E"))

summary(div_testE1)
anova(div_testE1)
```
Elite varieties Tuber yield seems to affected nothing. Insensitive to microbial diveristy. 



##Do post-hoc contrast

**Rick wants to split between Timepoints and Plot**

```{r}
div_testE1 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "E" & timepoint == "16S"))
div_testE2 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "E" & timepoint == "16S2"))
div_testA1 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "A" & timepoint == "16S"))
div_testA2 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "A" & timepoint == "16S2"))
div_testC1 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "C" & timepoint == "16S"))
div_testC2 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "C" & timepoint == "16S2"))
div_testSB1 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "SB" & timepoint == "16S"))
div_testSB2 <- lmer(Tubermass2 ~ Nutrient*Soil_type*diversity + (1|Variety), data = subset(metadata, Species == "SB" & timepoint == "16S2"))
```


***

#PCoA Analysis (Figure S5):

```{r}
ghsoilspecies_16S_PCA = subset(ghsoilspecies_16S_edited, timepoint == "16S2")
ghsoilpco = capscale(ghsoilspecies_16S_PCA[27:1504]~1, distance = "bray", data = ghsoilspecies_16S_PCA)
plot(ghsoilpco, display = "sites", type= "n", xlab = "MDS1(9.71%)", ylab = "MDS2(7.15%)")
points(ghsoilpco,display = "sites", col = as.numeric(ghsoilspecies_16S_PCA$Species), pch = 19)
ordiellipse(ghsoilpco, draw = "polygon",  ghsoilspecies_16S_PCA$Species, col = 1:4, border = 1:4, alpha = 63, label = F)
legend(x = "bottomleft", legend = levels(ghsoilspecies_16S_PCA$Species), col = 1:4, pch=20)
```

***


